Automating the Ontological Argument

### 2015-12-11 Papers/InProgress/IJCAI/paper.tex
IJCAI draft/plan added

\typeout{IJCAI-15 Instructions for Authors}

@@

\title{IJCAI--15 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
\author{Qiang Yang \\
Hong Kong University of Science and Technology\\
Hong Kong, China \\
pcchair15@ijcai.org}

@@

\begin{abstract}
  The {\it IJCAI--15 Proceedings} will be printed from electronic
  manuscripts submitted by the authors. The electronic manuscript will
  also be included in the online version of the proceedings. This paper
  provides the style instructions.
\end{abstract}

\section{Introduction}

The {\it IJCAI--15 Proceedings} will be printed from electronic
manuscripts submitted by the authors. These must be PDF ({\em Portable
Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.

\subsection{Length of Papers}

Each accepted full paper is allocated six pages in the conference 
proceedings, excluded references. References can take up to one page.
Up to two additional pages may be purchased at a price 
of \$275 per page for any accepted paper. However, all 
{\em submissions} must 
be a maximum of six pages, plus at most one for references, in length.

\subsection{Word Processing Software}

As detailed below, IJCAI has prepared and made available a set of
\LaTeX{} macros and a Microsoft Word template for use in formatting
your paper. If you are using some other word processing software (such
as WordPerfect, etc.), please follow the format instructions given
below and ensure that your final paper looks as much like this sample
as possible.

\section{Style and Format}

\LaTeX{} and Word style files that implement these instructions
can be retrieved electronically. (See Appendix~\ref{stylefiles} for
instructions on how to obtain these files.)

\subsection{Layout}

Print manuscripts two columns to a page, in the manner in which these
instructions are printed. The exact dimensions for pages are:
\begin{itemize}
\item left and right margins: .75$''$
\item column width: 3.375$''$
\item gap between columns: .25$''$
\item top margin---first page: 1.375$''$
\item top margin---other pages: .75$''$
\item bottom margin: 1.25$''$
\item column height---first page: 6.625$''$
\item column height---other pages: 9$''$
\end{itemize}

All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
A4-size paper, use the given top and left margins, column width,
height, and gap, and modify the bottom and right margins as necessary.

\subsection{Format of Electronic Manuscript}

For the production of the electronic manuscript, you must use Adobe's
{\em Portable Document Format} (PDF). A PDF file can be generated, for
instance, on Unix systems using {\tt ps2pdf} or on Windows systems
using Adobe's Distiller. There is also a website with free software
and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
\LaTeX2e{}, this is accomplished by putting
\begin{quote} 
\mbox{\tt $\backslash$usepackage\{times\}}
\end{quote}
in the preamble.\footnote{You may want also to use the package {\tt
latexsym}, which defines all symbols known from the old \LaTeX{}
version.}

Additionally, it is of utmost importance to specify the American {\bf
letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
formatting the paper. When working with {\tt dvips}, for instance, one
should specify {\tt -t letter}.

\subsection{Title and Author Information}

Center the title on the entire width of the page in a 14-point bold
font. Below it, center the author name(s) in a 12-point bold font, and
then center the address(es) in a 12-point regular font. Credit to a
sponsoring agency can appear on the first page as a footnote.

\subsubsection{Blind Review}

In order to make blind reviewing possible, authors must omit their
names and affiliations when submitting the paper for review. In place
of names and affiliations, provide a list of content areas. When
referring to one's own work, use the third person rather than the
first person. For example, say, ``Previously,
Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
that\ldots'' Try to avoid including any information in the body of the
paper or references that would identify the authors or their
institutions. Such information can be added to the final camera-ready
version for publication.

\subsection{Abstract}

Place the abstract at the beginning of the first column 3$''$ from the
top of the page, unless that does not leave enough room for the title
and author information. Use a slightly smaller width than in the body
of the paper. Head the abstract with ``Abstract'' centered above the
body of the abstract in a 12-point bold font. The body of the abstract
should be in the same font as the body of the paper.

The abstract should be a concise, one-paragraph summary describing the
general thesis and conclusion of your paper. A reader should be able
to learn the purpose of the paper and the reason for its importance
from the abstract. The abstract should be no more than 200 words long.

\subsection{Text}

The main body of the text immediately follows the abstract. Use
10-point type in a clear, readable font with 1-point leading (10 on
11).

Indent when starting a new paragraph, except after major headings.

\subsection{Headings and Sections}

When necessary, headings should be used to separate major sections of
your paper. (These instructions use many headings to demonstrate their
appearance; your paper should have fewer headings.)

\subsubsection{Section Headings}

Print section headings in 12-point bold type in the style shown in
these instructions. Leave a blank space of approximately 10 points
above and 4 points below section headings.  Number sections with
arabic numerals.

\subsubsection{Subsection Headings}

Print subsection headings in 11-point bold type. Leave a blank space
of approximately 8 points above and 3 points below subsection
headings. Number subsections with the section number and the
subsection number (in arabic numerals) separated by a
period.

\subsubsection{Subsubsection Headings}

Print subsubsection headings in 10-point bold type. Leave a blank
space of approximately 6 points above subsubsection headings. Do not
number subsubsections.

\subsubsection{Special Sections}

You may include an unnumbered acknowledgments section, including
acknowledgments of help from colleagues, financial support, and
permission to publish.

Any appendices directly follow the text and look like sections, except
that they are numbered with capital letters instead of arabic
numerals.

The references section is headed ``References,'' printed in the same
style as a section heading but without a number. A sample list of
references is given at the end of these instructions. Use a consistent
format for references, such as that provided by Bib\TeX{}. The reference
list should not include unpublished work.

\subsection{Citations}

Citations within the text should include the author's last name and
the year of publication, for example~\cite{gottlob:nonmon}.  Append
lowercase letters to the year in cases of ambiguity.  Treat multiple
authors as in the following examples:~\cite{abelson-et-al:scheme}
or~\cite{bgf:Lixto} (for more than two authors) and
\cite{brachman-schmolze:kl-one} (for two authors).  If the author
portion of a citation is obvious, omit it, e.g.,
Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
\nocite{abelson-et-al:scheme}
\nocite{bgf:Lixto}
\nocite{brachman-schmolze:kl-one}
\nocite{gottlob:nonmon}
\nocite{gls:hypertrees}
\nocite{levesque:functional-foundations}
\nocite{levesque:belief}
\nocite{nebel:jair-2000}

\subsection{Footnotes}

Place footnotes at the bottom of the page in a 9-point font.  Refer to
them with superscript numbers.\footnote{This is how your footnotes
should appear.} Separate them from the text by a short
line.\footnote{Note the line separating these footnotes from the
text.} Avoid footnotes as much as possible; they interrupt the flow of
the text.

\section{Illustrations}

Place all illustrations (figures, drawings, tables, and photographs)
throughout the paper at the places where they are first discussed,
rather than at the end of the paper. If placed at the bottom or top of
a page, illustrations may run across both columns.

Illustrations must be rendered electronically or scanned and placed
directly in your document. All illustrations should be in black and
white, as color illustrations may cause problems. Line weights should
be 1/2-point or thicker. Avoid screens and superimposing type on
patterns as these effects may not reproduce well.

Number illustrations sequentially. Use references of the following
form: Figure 1, Table 2, etc. Place illustration numbers and captions
under illustrations. Leave a margin of 1/4-inch around the area
covered by the illustration and caption.  Use 9-point type for
captions, labels, and other text in illustrations.

@@

The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
files that implement them was supported by Schlumberger Palo Alto
Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
Preparation of the Microsoft Word file was supported by IJCAI.  An
early version of this document was created by Shirley Jowell and Peter
F. Patel-Schneider.  It was subsequently modified by Jennifer
Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
These instructions are the same as the ones for IJCAI--05, prepared by
Kurt Steinkraus, Massachusetts Institute of Technology, Computer
Science and Artificial Intelligence Lab.

>>>>

\typeout{Automating the Ontological Argument}

@@

\title{Automating the Ontological Argument}
\author{Christoph Benzmüller\thanks{German Research Foundation DFG \ldots} and Bruno Woltzenlogel Paleo}
\author{}

@@

\begin{abstract}
  Higher-order automated theorem provers have revealed some
  philosophically profound new knowledge in metaphysics.
\end{abstract}

\section{Introduction}
G\"{o}del's ontological argument for the existence of God is frequently
presented as a masterpiece argument in metaphysics. A huge body of
literature is available which discusses G\"{o}del's original work and/or
subsequent variants of it.

This paper presents an in depth analysis of G\"{o}del's original version
of the ontological argument.

The analysis has been conducted with automated theorem provers for
classical higher-order logic (HOL). However, since higher-order modal
logic is actually required for the encoding of G\"{o}del's argument, we
utilize an semantic embedding of HOML in HOL.

First applications of theorem proving technology in metaphysics
appeared in 199? in the work Zalta and Oppenheimer.  Further work also
includes Rushby \ldots. Different to what is presented here, none of
these previous formalises and automates variants of HOML, which is
however crucial for G\"{o}del's ontological argument.

The work presented here substantially extends previous work \cite{}.  
The novel contributions of this paper include:

\begin{itemize}
\item Detailed analysis of the inconsistency of G\"{o}del's axioms. The
inconsistency was detected before, but for a long time we were not
able to extract an intuitive argument from the proof of the ATPs.
Further experiments now led to discovery of a surprisingly accessible
understanding of this inconsistency which is philosophically profound
and which has not been presented yet in the respective literature.

\item Fully automated proof of Scott's theorem T3 from the axioms, that
   is, without intermediate argumentation steps as used in previous
   work.

\item Direct verification of modal collapse with Meson in Isabelle/HOL,
which did not work for the previous.

\item On a technical level we show that the novel embedding of S5 is more
    effective than the previous one.
\end{itemize}

These novel contributions are presented in Sections ? and ?.

In Section ?, novel and previous findings are summarised and discussed 
with regard to their relevance in metaphysics. In this section also 
some further details and improvements on earlier 
findings are presented.

Section ? concludes the paper.

\section{Variants of the Ontological Argument}
Short history, Anselm, Leibniz, etc.
\subsection{G\"{o}del's Proof Script}
\subsection{Scott's Variant}
\subsection{Recent Variants}

\section{Automating HOML in HOL}
\subsection{Outline of the Embedding}
\subsection{Improved Embedding for S5}

\section{Intuitive Inconsistency Argument}
\subsection{LEO-II's Inaccessible Inconsistency Proof}
Tell the story (here or earlier). Show parts of LEO proof.  Inaccessible, but relevant knowledge
is contained.

\subsection{Argument Reconstruction in Isabelle}
Present the argument informal and in Isabelle. Easy to understand in 
in KB and KT, slightly harder in K.

\section{Full Automation of Scott's Proof}
Within the modified logic S5 Scott's proof can be fully automated. The
intermediate argumentation steps as not needed anymore to support the
prover.

In the modified S5 setting also further results can be automatically verified, which 
failed in previous work: verification of  modal collapse in
Isabelle/HOL with Meson.

\section{Conclusion}

Good improvement (technical sense), Sledgehammer can still be improved
(LEO-II finds inconsistency when call directly in THF Syntax, but not
from within Isabelle). Without independent experiments directly in THF 
the inconsistency would thus not have been detected.

@@

Will be added at a later point.

%German Research Foundation DFG, Chad Brown
